### 2015-02-26 Documentation/i.palachev/Thesis/Lomonosov2015/palachev.tex
[DOC] lomonosov2015: template, title and references
Signed-off-by: Ilya Palachev <iliyapalachev@gmail.com>

\documentclass[usePics]{lomonosov}

\begin{thesis}

\Title{О методах восстановления геометрических тел по измерениям опорной функции
и их приложении к задаче восстановления тел по теневым
контурам}{{Палачев\,И.\,А.}}

\Author{Палачев~Илья~Александрович}{Аспирант}{Механико-математический
факультет МГУ имени М.\,В.\,Ломоносова}{Москва}{Россия}{palachev.ilya@yandex.ru}

\begin{references}
\Source \ENGLISH{Prince\,J.\,L., Willsky\,A.\,S. Reconstructing convex sets
        from support line measurements // IEEE Transactions on Pattern Analysis
        and Machine Intelligence, 12(4):377-389, 1990.}

\Source \ENGLISH{Lele\,A.\,S., Kulkarni\,S.\,R., and Willsky\,A.\,S.
        Convex-polygon estimation from support-line measurements and
        applications to target reconstruction from laser-radar data // Journal
        of the Optical Society America A, 9(10):1693-1714, Oct 1992.}

\Source \ENGLISH{Karl\,W.\,C., Kulkarni\,S.\,R., Verghese\,G.\,C.,
        Willsky\,A.\,S. Local tests for consistency of support hyperplane data
        // Journal of Mathematical Imaging and Vision, 6(2-3):249-267, 1996.}

\Source \ENGLISH{Gregor\,J., Rannou\,F.\,R. 3D Support Function Estimation
        and Application for Projection MRI // International Journal for
        Imaging Systems and Technology, 12:43-50, 2002.}

\Source \ENGLISH{Gardner\,R.\,J., Kiderlen\,M. A new algorithm for 3D
        reconstruction from support functions // IEEE Transactions on Pattern
        Analysis and Machine Intelligence, 31(3):556-562, 2009.}
\end{references}

\end{thesis}